Untwisting the Tetrahelix (v0.6)

### 2017-02-27 doc/TetrahelixUntwist.tex
attempting to improve proofs

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newtheorem{proofsketch}{Proof Sketech}

@@

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{d_{bc} 2 \pi }{\rho_{bc}} \\  
  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational, a BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch. For
example, a slight irregularity that led to a pitch of precisely 10
tetrahedra in one revolution would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch.

@@

We use the term \emph{tetrahelix} to mean any structure made of
vertices and edges which is isomorphic to the BC helix, in which the
vertices lie on three helices, no matter what lengths the edges take
on. One could consider various definitions of optimality for a
tetrahelix, but the must useful to an enginner is to minimize the
maximum difference between any two edges.
Define the \emph{projected incircle radius} of a tetrahelix to be
the radius of the incircle of all point on the tetrahelix projected
down the $z$ axis into the $xy$ plane.
We call a tetrahelix
\emph{member-length minimax optimal} (or just \emph{optimal}) if there is not tetrahelix of the
same project incircle radius and pitch with a smaller maximum edge length difference.

Consider a tetrahelix isomporphic to an BC helix of unbounded
extent. If all three rails do not have same pitch, there is an edge of
unbounded length, so it is not optimal. So we are justified in talking about the pitch of
the tetrahelix, even though we think of it as three
helices. Similarly, if the axes are not parallel, there is an edge of
unbounded length in the structure, so it is not optimal. So all three
rails have the same rail length.

We wish to show that the three helices optimal tetrahelix also have
coincident axes.
Consider any tetrahelix in which the axes of the helices are parallel
to $z$ but not coincident, but in which all vertices lie on or outside
a cylinder of radius $r$ centered on the $z$ axis.
Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Consider the projection along the $z$ axis of all vertices into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or an polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definitions lies outside the
circle of radius $r$ in the $xy$ plane.
It is easiest to consider helices of zero curvature which only have one point
as a special case. 
For one rail consider a point in this projected figure furthest from the midpoint of the
the two rails. The lengths to the furthest point on the other two rails will
be decreased by moving the axis of that rail closer to the midpoint of the
other two. Such motion 
One of these segements
will be longer and will be made shorter by moving that helix closer to
the midpoint of the other two in the $xy$ plane. Even if the rail
angle is $\pi$, there will be two vertices for each rail, each of
which is connected to both vertices of the other two rails. At lesast
one of these lines is a longest line which will be made shorter by
moving that helix in the $xy$ plane closer to the midpoint of the
other two.

For the special case of degenerate helix where we do not know the angle
of the points from the axis, we can see that we minimize the maximum lengths
when the projected vertices form an equilateral triangle around the $z$ axis at
length $6r/sqrt{3}$.
This is a equivalent to some let of helices which do have their axes on
the $z$ axis, so without loss of generality we treat these degenerate
helices as if they have their axis at the $z$-axis.

So any any optimal tetrahelix with a rail-angle of greater than $0$,
that is, with any curvature, will have conincident axes.

We call a tetrahelix with $0$ curvature a \emph{tetrabeam}, and
consider it as an important special case and an endpoint of a
continuum of tetrahelices of decreasing curvature.

 Now that we have coincident axes, the same pitch, we can go on to
 the hard proof about where they occur.

@@

\footnote{Another interesting but non-optimal solution is derived by setting (one-hop + two-hop)/2 = 1,  occurs at $r = \sqrt{35}/4$,
which produces three length classes of $11/12, 12/12, 13/12$!}

@@

@@

>>>>

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}

@@

Since:
\[ \frac{2 \pi}{\rho_{bc}} \approx 10.16
\]
We can see that there are approximately $10.16$ red, blue or yellow tetrahedra on one rail in a single revolution.
The pitch of the Boerdijk--Coxeter helix of edge length $1$ is the length of three tetrahedra times this number:
\begin{align*}
  &= \frac{d_{bc} 2 \pi }{\rho_{bc}} \\  
  &= \frac{3 \cdot h_{bc} 2 \pi }{\rho_{bc}} \\
  &= \frac{3  \sqrt{\frac{2}{5}}  \pi}{\rho_{bc}} \\
  &\approx 9.64 \\
\end{align*}
The pitch is less than the number of tetrahedra because the tetrahedra
are not lined up perfectly.  It is a famous and interesting result
that the pitch is irrational, a BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of deveoping a rational period by arranging solid tetrahedra by relaxing the face-to-face matching
has been explored\cite{sadler2013periodic}. In this paper
we develop slightly irregular edge lengths that support, for example, a pitch of precisely 10
tetrahedra in one revolution which would allow an architect to design a
column having a basis and a capital in the same relation to the
tetrahedra they touch at the bottom and top of the column.

@@

We use the term \emph{tetrahelix} to mean any structure made of
vertices and edges which is isomorphic to the BC helix, in which the
vertices lie on three helices. We further demand that all edge lengths
be finite, as we are only interested in physically constructable tetrahelices.
By isomporphic we mean there is a one-to-one mapping between both
vertices and edges in the two tetrahelices.
One could consider various definitions of optimality for a
tetrahelix, but the must useful to an enginner is to minimize the
maximum difference between any two edges.

If all three rails do not have the same pitch, there is an edge of
unbounded length, so it is not optimal. So we are justified in talking about the
\emph{pitch} of 
the tetrahelix as the pitch of its rail helices, even though there are
three such helices.

Similarly, if the axes are not parallel, there is an edge of
unbounded length in the structure, so we exclude such a strcturue.

Since the axes are parallel, we may define the \emph{inradius} of a tetrahelix to be the radius of the largest
``in cylinder'': a
cylinder parallel to this axis contained within circumscribing cylinder which is penetrated by no edge.
Define an \emph{minimax edge-length optimal tetrahelix} or just an
\emph{optimal tetrahelix} to be a tetrahelix for which there exists
no other tetrahelix of the same inradius and pitch with a lower maximum
difference between its edge lengths.

We wish to show that in an optimal tetrahelix, all vertices lie on the cylinder
of radius $r$, regardless of where they lie on the $z$-axis.

Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Our tool for thinking about this is to project out the $z$ dimension to form
a two-dimensional figure of nodes and non-rail edges.
Consider the projection along the $z$ axis of all vertices and non-rail edges into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or an polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definitions lies outside the
circle of radius $r$ in the $xy$ plane.

\begin{theorem}
  Any optimal tetrahelix has all vertices on a single cylinder.
\end{theorem}

\begin{proof}

Case 1: Suppose that $\rho$ is zero. Then for any given inradius, an equilateral
triangle is the minimax solution for all non-rail edges. Since there is only
one rail-edge length, this is the minimax solution for the entire set. Since
the projection is an equilateral triangle lies on a circle, all points lie on
a cylinder.

Case 2: Suppose that $\rho$ is positive but less than $\pi$.
In this case each rail helix has
curavature and places points on both sides of any line through the origin
of the $xy$ plane (or both coincident on such a line.)

We first show that the inradius is touched by one segment from each pair
of rails.

\begin{figure}[H]
  \label{untouchingrailfig}
     \centering
     \includegraphics[width=0.7\textwidth]{figures/UntouchingRail.png}
     \caption{Untouching Rail}
\end{figure}

Figure\ref{untouchingrailfig} depicts a projection of
depicting only the point $P$, the point $P'$ which is an adjacent vertex on the same rail,
and the intermediary points $R$ and $Q$ on the other two rails.

Supoose there is a rail $P$ which does not have a segment touching the incircle
at all in the projection, as depicted in 
Then a segment connecting the other two $Q$ and $R$ is a longest
rail, because any chord touching an incircle is longer than any which does not.
These two rails ($Q$ and $R$) can be moved closer to each other, and further away from $P$, to form a better minimax solution
by shortening the longest rail, until one of the segments from $P$ the previously
untouching rail touches one of the incircle. By induction then all rails have
at least one segement touch each vertex on the rail that is tangent to the
the inradius in an optimal tetrahelix.

Now suppose that you attempt to move the axis of one of the rails. Because
$\rho > 0 $ and $\rho$ is not a multiple of $\pi$, there is some vertex on the two sides of any diameter
halving our circle. Moving the axis of a tetrahelix lengthens a longest line while ``pinching'' the
inradius on the other side, so it is not optimal. Since the axes of these helices are the
same and they have the same curvature and pitch, the points defined by the helices all lie
on a cylinder.

\end{proof}

Now that we have show that any optimal tetrahelix vertices that
are on helices of the same axes and pitch, we see that the vertices 
of any optimal tetrahelix will lie on a cylinder, or a circle when axis dimension
is projected out. Therefore is is reasonable to now speak of the \emph{radius}
of a tetrahelix as the radius of the cylinder, and the notion
of the projected incircle radius is now obsolete.

Now that we have coincident axes, the same pitch, we can go on to
 the hard proof about where they occur along the $z$-axis.

@@

\footnote{Another interesting but non-optimal solution is derived by setting (one-hop + two-hop)/2 = 1,  occurs at $r = \sqrt{35}/4$,
which produces three length classes of $11/12, 12/12, 13/12$}

@@

  Consider any tetrahelix in which the axes of the helices are parallel
to $z$ but not coincident, but in which all vertices lie on or outside
a cylinder of radius $r$ centered on the $z$ axis.
Since all three rails have the same rail length, no matter how we
move the rails in the $xy$ plane if we shorten the $xy$ distance between
vertices we shorten the total distance.
Consider the projection along the $z$ axis of all vertices and non-rail edges into the $xy$ plane, which will be
a figure of dots and connecting segments in the $xy$ plane. The convex
hull for any one helix projection will be a circle (if its pitch is
irrational) or an polygon if rational, or a point if the helix has
zero curvature. Each of these figures by definitions lies outside the
circle of radius $r$ in the $xy$ plane.
It is easiest to consider helices of zero curvature which only have one point
as a special case. 
For one rail consider a point in this projected figure furthest from the midpoint of the
the two rails. The lengths to the furthest point on the other two rails will
be decreased by moving the axis of that rail closer to the midpoint of the
other two.
One of these segements
will be longer and will be made shorter by moving that helix closer to
the midpoint of the other two in the $xy$ plane. Even if the rail
angle is $\pi$, there will be two vertices for each rail, each of
which is connected to both vertices of the other two rails. At lesast
one of these lines is a longest line which will be made shorter by
moving that helix in the $xy$ plane closer to the midpoint of the
other two.

For the special case of degenerate helix where we do not know the angle
of the points from the axis, we can see that we minimize the maximum lengths
when the projected vertices form an equilateral triangle around the $z$ axis at
length $6r/\sqrt{3}$.
This is a equivalent to some set of helices which do have their axes on
the $z$ axis, so without loss of generality we treat these degenerate
helices as if they have their axis at the $z$-axis.

So any any optimal tetrahelix with a rail-angle of greater than $0$,
that is, with any curvature, will have conincident axes, or are indistinguishable
from such a set.

Possible addtional refrerences and submissions:

Journal of applied geometry seems like a good fit:

http://epubs.siam.org/toc/sjaabq/1/1

According to there editorial board, they welcome mathematcal software and
online resources such as I provie.

This one has a nice editing structure: Metric Geometry is a nice subset:

http://discreteanalysisjournal.com/for-authors

@@

This is unrelated, but well worth reading:

https://arxiv.org/pdf/1702.06388.pdf
